### 2020-02-29 thesis/content/chapter-related-work.tex
fixed font issues

The \acf{wl} algorithm~\cite{Weisfeiler1968}\cite{Cai1992} characterizes a graph $G$ by assigning discrete labels $c \in \mathcal{C}$, called \textit{colors}, to vertex $k$-tuples $(v_1, \dots, v_k) \in \mathcal{V}_G^k$, where $k \in \mathbb{N}$ is the freely choosable \textit{\ac{wl}-dimension}.
A mapping $\chi_{G, k}: \mathcal{V}_G^k \to \mathcal{C}$ is called a \textit{$k$-coloring} of $G$.
\begin{defn}
	A coloring $\chi'$ \textit{refines} $\chi$ ($\chi' \preceq \chi$) iff.\ $\forall a, b \in \mathcal{V}_G^k: \chi(a) \neq \chi(b) \rightarrow \chi'(a) \neq \chi'(b)$, i.e.\ $\chi'$ distinguishes at least those tuples that are distinguished by $\chi$.
\end{defn}
\begin{defn}
	Two colorings $\chi$ and $\chi'$ are \textit{equivalent} ($\chi \equiv \chi'$) iff.\ $\chi \preceq \chi' \land \chi' \preceq \chi$, i.e.\ $\chi$ is identical to $\chi'$ up to color substitutions.
\end{defn}
The $k$-dimensional \ac{wl} algorithm ($k$-\acs{wl}) works by iteratively refining $k$-colorings $\chi_{G, k}^{(0)} \succeq \chi_{G, k}^{(1)} \succeq \dots$ of a given graph $G$ until the convergence criterion $\chi_{G, k}^{(i)} \equiv \chi_{G, k}^{(i+1)}$ is satisfied.
We denote the final, maximally refined $k$-\acs{wl} coloring with $\chi^{*}_{G, k}$.
\begin{defn}
	The color distribution $\mathit{dist}_{\chi_{G, k}}: \mathcal{C} \to \mathbb{N}$ of a $k$-coloring $\chi_{G, k}$ counts each color $c \in \mathcal{C}$ in the coloring, i.e.\ $\mathit{dist}_{\chi_{G, k}}(c) \coloneqq \left|\left\{ v \in \mathcal{V}_G^k\, |\, \chi_{G, k}(v) = c \right\}\right|$. % chktex 21
\end{defn}
\begin{defn}\label{defn:related:wl-distinguishable}
	Two graphs $G$ and $H$ are are $k$-\acs{wl} \textit{distinguishable} ($G \mathrel{{\not\simeq}_k} H$) iff.\ there exists a color $c \in \mathcal{C}$ s.t.\ $\mathit{dist}_{\chi^{*}_{G, k}}(c) \neq \mathit{dist}_{\chi^{*}_{H, k}}(c)$.
\end{defn}
As we will see, the way in which \ac{wl} colorings are refined is vertex order invariant;
thus any difference in the final coloring of two graphs always implies the non-isomorphism of the colored graphs, i.e.\ $G \mathrel{{\not\simeq}_k} H \implies G \not\simeq H$.
The opposite does however not necessarily hold;
two $k$-\acs{wl} indistinguishable graphs are not always isomorphic, i.e.\ $\exists\, G, H: G \mathrel{{\simeq}_k} H \land G \not\simeq H$. % chktex 21

@@

For $k = 1$ \citet{Babai1980} have shown that two randomly selected non-isomorphic graphs $G \not\simeq H$ of size $n$ are 1-\acs{wl} indistinguishable with probability $n^{-\nicefrac{1}{7}}$.
Thus 1-\acs{wl} is already able to distinguish most graphs; it fails however to distinguish any pair of unlabeled $d$-regular graphs of the same size~\cite[cor.~1.8.5]{Immerman1990}.
\begin{defn}
	A graph $G$ is called $d$-regular ($\mathit{rg}_d(G)$) iff.\ $\forall v \in \mathcal{V}_G: {|\Gamma_G(v)|} = d$.
\end{defn}
We have already seen this in \cref{fig:related:wl1-problem} since the ``six-cycle'' and the ``two-triangles'' graphs are in fact both 2-regular and of size 6.
This restriction alone is typically not an issue in the context of \ac{gcr} though, as molecular structures or social interaction graphs for example are rarely perfectly regular.
A more relevant restriction of 1-\acs{wl} is the fact that it is unable to detect cycles of length $m \geq 3$.
\begin{defn}\label{defn:related:wl-compute}
	$k$-\acs{wl} computes a function $f: \mathcal{G} \to Y$ iff.\ that function can be expressed as $f(G) = g(\mathit{dist}_{\chi_{G, k}^{*}})$ via some function $g: (\mathcal{C} \to \mathbb{N}) \to Y$.
\end{defn}
\begin{defn}\label{defn:related:wl-count-detect}
	$k$-\acs{wl} counts a certain subgraph $S$ iff.\ it computes $\mathit{count}_S(G) \coloneqq \left|\{ \hat{v}\, |\, v \in \mathcal{V}_G^{*} \land G[v] \equiv S \}\right|$ (see~\cref{defn:related:ordered-subgraph}).
	Similarly $k$-\acs{wl} detects a subgraph $S$ iff.\ it computes $\mathit{contain}_S(G) \coloneqq \mathbbm{1}[\mathit{count}_S(G) > 0]$, with $\mathbbm{1}$ denoting the indicator function.
\end{defn}
To see why 1-\acs{wl} is unable to detect $m$-cycles in graphs, note that \cref{fig:related:wl1-problem} already contradicts the positive statement for $m = 3$;
this counterexample can be trivially generalized to all larger $m$ by replacing the ``six-cycle'' graph with a ``$(2m)$-cycle-graph'' graph and the ``two-triangles'' graph with a ``two-$m$-cycles'' graph which preserves the 2-regularity and therefore the 1-\acs{wl} indistinguishability.

@@

The \ac{ft} of a graph signal $x$ then is $\hat{x}[i] = \langle u_{G,i}, x \rangle$ with $u_G = {\{ u_{G,i} \}}_{i=1}^{n}$ being the eigenvectors of $L_G$, s.t.\ $x = \sum_{i=1}^n \hat{x}[i] u_{G,i}$.
Note that $L_G$ is guaranteed to only have real eigenvectors $u_G$ and eigenvalues $\lambda_G$ due to the fact that we only consider undirected graphs with symmetric adjacency matrices $A_G$.
By convention we assume that the eigenvalues are ordered ascendingly: $\lambda_{G,0} \leq \cdots \leq \lambda_{G,n}$.
Those eigenvalues are called the \textit{spectrum of $G$} and they are a vertex-permutation-invariant graph characterization.
Intuitively those eigenvalues describe the connectivity between different parts of the graph.
The initial eigenvalues represent connectivity at a general, coarse level while the last eigenvalues represent the connectivity between finer substructures.
\Cref{fig:related:graph-fourier} illustrates this idea.
A more concrete example for the relation between a graph's structure and its spectrum is that a graph with $m$ connected components has exactly $m$ zero eigenvalues, i.e.\ $\lambda_{G,1} \cdots = \lambda_{G,m} = 0$.
\begin{figure}[ht]
	\centering
	\includegraphics[width=\linewidth]{gfx/related-work/graph-fourier.pdf}
	\caption[Comparison between the basis vectors of the real domain \acl{ft} and the graph \acl{ft}.]{
		Comparison between the basis functions/vectors of the real domain \ac{ft} and the graph \ac{ft}.
		For the eigenfunctions on the upper half only the real cosine components of the complex exponentials are shown.
	}\label{fig:related:graph-fourier}
\end{figure}

>>>>

The \acf{wl} algorithm~\cite{Weisfeiler1968}\cite{Cai1992} characterizes a graph $G$ by assigning discrete labels $c \in \mathcal{C}$, called \textit{colors}, to vertex $k$-tuples $(v_1, \dots, v_k) \in \mathcal{V}_G^k$, where $k \in \mathbb{N}$ is the freely choosable \textit{\ac{wl}-dimension}.
A mapping $\chi_{G, k}: \mathcal{V}_G^k \to \mathcal{C}$ is called a \textit{$k$-coloring} of $G$.
\begin{defn}
	A coloring $\chi'$ \textit{refines} $\chi$ ($\chi' \preceq \chi$) iff.\ $\forall a, b \in \mathcal{V}_G^k: \chi(a) \neq \chi(b) \rightarrow \chi'(a) \neq \chi'(b)$, i.e.\ $\chi'$ distinguishes at least those tuples that are distinguished by $\chi$.
\end{defn}
\begin{defn}
	Two colorings $\chi$ and $\chi'$ are \textit{equivalent} ($\chi \equiv \chi'$) iff.\ $\chi \preceq \chi' \land \chi' \preceq \chi$, i.e.\ $\chi$ is identical to $\chi'$ up to color substitutions.
\end{defn}
The $k$-dimensional \ac{wl} algorithm ($k$-\acs{wl}) works by iteratively refining $k$-colorings $\chi_{G, k}^{(0)} \succeq \chi_{G, k}^{(1)} \succeq \dots$ of a given graph $G$ until the convergence criterion $\chi_{G, k}^{(i)} \equiv \chi_{G, k}^{(i+1)}$ is satisfied.
We denote the final, maximally refined $k$-\acs{wl} coloring with $\chi^{*}_{G, k}$.
\begin{defn}
	The color distribution $\mathit{dist}_{\chi_{G, k}}: \mathcal{C} \to \mathbb{N}$ of a $k$-coloring $\chi_{G, k}$ counts each color $c \in \mathcal{C}$ in the coloring, i.e.\ $\mathit{dist}_{\chi_{G, k}}(c) \coloneqq \left|\left\{ v \in \mathcal{V}_G^k\, |\, \chi_{G, k}(v) = c \right\}\right|$. % chktex 21
\end{defn}
\begin{defn}\label{defn:related:wl-distinguishable}
	Two graphs $G$ and $H$ are $k$-\acs{wl} \textit{distinguishable} ($G \mathrel{{\not\simeq}_k} H$) iff.\ there exists a color $c \in \mathcal{C}$ s.t.\ $\mathit{dist}_{\chi^{*}_{G, k}}(c) \neq \mathit{dist}_{\chi^{*}_{H, k}}(c)$.
\end{defn}
As we will see, the way in which \ac{wl} colorings are refined is vertex order invariant;
thus any difference in the final coloring of two graphs always implies the non-isomorphism of the colored graphs, i.e.\ $G \mathrel{{\not\simeq}_k} H \implies G \not\simeq H$.
The opposite does however not necessarily hold;
two $k$-\acs{wl} indistinguishable graphs are not always isomorphic, i.e.\ $\exists\, G, H: G \mathrel{{\simeq}_k} H \land G \not\simeq H$. % chktex 21

@@

For $k = 1$ \citet{Babai1980} have shown that two randomly selected non-isomorphic graphs $G \not\simeq H$ of size $n$ are 1-\acs{wl} indistinguishable with probability $n^{-\nicefrac{1}{7}}$.
Thus 1-\acs{wl} is already able to distinguish most graphs; it fails however to distinguish any pair of unlabeled $d$-regular graphs of the same size~\cite[cor.~1.8.5]{Immerman1990}.
\begin{defn}
	A graph $G$ is called \textit{$d$-regular} ($\mathit{rg}_d(G)$) iff.\ $\forall v \in \mathcal{V}_G: {|\Gamma_G(v)|} = d$.
\end{defn}
We have already seen this in \cref{fig:related:wl1-problem} since the ``six-cycle'' and the ``two-triangles'' graphs are in fact both 2-regular and of size 6.
This restriction alone is typically not an issue in the context of \ac{gcr} though, as molecular structures or social interaction graphs for example are rarely perfectly regular.
A more relevant restriction of 1-\acs{wl} is the fact that it is unable to detect cycles of length $m \geq 3$.
\begin{defn}\label{defn:related:wl-compute}
	$k$-\acs{wl} \textit{computes} a function $f: \mathcal{G} \to Y$ iff.\ that function can be expressed as $f(G) = g(\mathit{dist}_{\chi_{G, k}^{*}})$ via some function $g: (\mathcal{C} \to \mathbb{N}) \to Y$.
\end{defn}
\begin{defn}\label{defn:related:wl-count-detect}
	$k$-\acs{wl} \textit{counts} a certain subgraph $S$ iff.\ it computes $\mathit{count}_S(G) \coloneqq \left|\{ \hat{v}\, |\, v \in \mathcal{V}_G^{*} \land G[v] \equiv S \}\right|$ (see~\cref{defn:related:ordered-subgraph}).
	Similarly $k$-\acs{wl} \textit{detects} a subgraph $S$ iff.\ it computes $\mathit{contain}_S(G) \coloneqq \mathbbm{1}[\mathit{count}_S(G) > 0]$, with $\mathbbm{1}$ denoting the indicator function.
\end{defn}
To see why 1-\acs{wl} is unable to detect $m$-cycles in graphs, note that \cref{fig:related:wl1-problem} already contradicts the positive statement for $m = 3$;
this counterexample can be trivially generalized to all larger $m$ by replacing the ``six-cycle'' graph with a ``$(2m)$-cycle-graph'' graph and the ``two-triangles'' graph with a ``two-$m$-cycles'' graph which preserves the 2-regularity and therefore the 1-\acs{wl} indistinguishability.

@@

The \ac{ft} of a graph signal $x$ then is $\hat{x}[i] = \langle u_{G,i}, x \rangle$ with $u_G = {\{ u_{G,i} \}}_{i=1}^{n}$ being the eigenvectors of $L_G$, s.t.\ $x = \sum_{i=1}^n \hat{x}[i] u_{G,i}$.
Note that $L_G$ is guaranteed to only have real eigenvectors $u_G$ and eigenvalues $\lambda_G$ due to the fact that we only consider undirected graphs with symmetric adjacency matrices $A_G$.
By convention we assume that the eigenvalues are ordered ascendingly: $\lambda_{G,0} \leq \cdots \leq \lambda_{G,n}$.
Those eigenvalues are called the \textit{spectrum of $G$} and they are a vertex-permutation-invariant graph characterization.
Intuitively those eigenvalues describe the connectivity between different parts of the graph.
The initial eigenvalues represent connectivity at a general, coarse level while the last eigenvalues represent the connectivity between finer substructures.
\Cref{fig:related:graph-fourier} illustrates this idea.
A more concrete example for the relation between a graph's structure and its spectrum is that a graph with $m$ connected components has exactly $m$ zero eigenvalues, i.e.\ $\lambda_{G,1} \cdots = \lambda_{G,m} = 0$.
\begin{figure}[ht]
	\centering
	\includegraphics[width=\linewidth]{gfx/related-work/graph-fourier.pdf}
	\caption[Comparison between the basis vectors of the real domain \acl{ft} and the graph \acl{ft}.]{
		Comparison between the basis functions/vectors of the real domain \ac{ft} and the graph \ac{ft}.
		For the eigenfunctions on the upper half only the real cosine components of the complex exponentials are shown.
		\source[based on]{Shuman2013}
	}\label{fig:related:graph-fourier}
\end{figure}
